The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-06-30 oblique.tex
updated bib

These potential implications of blazar heating rely on an understanding of the linear and nonlinear physics of these plasma instabilities.  Recent work in this area has been inconclusive.  For instance, \citet{Miniati+12} argued that these instabilities are physically irrelevant for the cooling of these pair beams because they would saturate at a very low level due to nonlinear Landau damping (NLD).  However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.  In addition, \citet{Siron+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.  It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g. Mohamad's paper).

@@

Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
\begin{equation}\label{eq:dispersion relation reactive}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
\end{equation}
For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{BCP}.

@@

The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.  However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.  However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.  First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.  This is a part of ongoing work and will be presented in a future publication. Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.  Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).  Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{Puchwein+2011}.

>>>>

These potential implications of blazar heating rely on an understanding of the linear and nonlinear physics of these plasma instabilities.  Recent work in this area has been inconclusive.  For instance, \citet{Miniati+12} argued that these instabilities are physically irrelevant for the cooling of these pair beams because they would saturate at a very low level due to nonlinear Landau damping (NLD).  However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.  In addition, \citet{Sironi+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.  It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g. Mohamad's paper).

@@

Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
\begin{equation}\label{eq:dispersion relation reactive}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
\end{equation}
For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{paperI}.

@@

The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.  However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.  However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.  First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.  This is a part of ongoing work and will be presented in a future publication. Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.  Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).  Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
